NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-06-21 Manuscript/manuscriptPS.tex
Update text

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPS.eps}
  \caption{\label{HGorderparametersPCvsPS}
    Changes of PC (left panel) and PS (right panel) headgroup order parameters
    from POPC:POPS mixtures with increasing amount of POPS.
    Experimental results of POPC are taken from Ref. \citenum{scherer87}
    (signs are determined as discussed in \cite{botan15,ollila16}).
    Experimental values for POPS in pure bilayer and in mixture are  
    measured in this work and in Ref. \citenum{roux90} at 298K, respectively.
    For POPC (left panel) the zero point of y-axis is set to the value of pure bilayer.
    For $\beta$-carbon of POPS (right panel, top) the zero point of y-axis is
    set to the value from POPC:POPS (5:1) mixture.
    For $\alpha$-carbon of POPS (right panel, bottom) the y-axis is tranferred
    with the same value for both order parameters such that the lower order
    parameter value from POPC:POPS (5:1) mixture is at zero
    to correctly illustrate the significant forking.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.} \\
  \todo{Simulation of pure POPC at 298K with Lipid14 would be useful for this plot (only at 303 K is available from NMRlipids I)} \\
  \todo{MacRog simulations of pure POPS with potassium counterions only would be useful for this and other plots.}
\end{figure*}

Also the headgroup order parameters of POPS mixed with varying amounts of POPC
from simulations and experiments \cite{borle85,roux90} are shown in Fig. \ref{HGorderparametersPCvsPS}.
The $\beta$-carbon order parameter of POPS slightly decreases and positive 
order parameter of $\alpha$-carbon slightly increases with increasing
amount of PS lipids. This may indicate increasing order of the headgroup.
\todo{The existing literature about this should be mentioned/discussed.}
It should be, however, noted that the experimental data for pure POPS and
POPC/POPS mixture come from different experimental sets, 13C NMR in this work
and 2H NMR from Ref. \citenum{roux90}, respectively. Therefore the accuracy
of the order parameter change is not as high as typically in the measurements
of order parameter changes, see discussion about qualitative and quantitative 
accuracy in Ref. \citenum{ollila16}. The changes of PS headgroup order parameters
are not reproduced by the tested simulation models. The $\beta$-carbon order parameter
increase with increasing amount of PS in CHARMM36 and MacRog simulations in contrast
to the experimental data. The smaller $\alpha$-carbon order parameter increase in both
simulation models with increasing amount of PS, while it is almost unchanged in experiments.
The larger $\alpha$-carbon order parameter increase in MacRog and decrease in CHARMM36
with increasing amount of PS, both model exhibiting a poor agreement with experiments. 
Signifincant improvement in the MD simulation models are needed to interpret the
PS headgroup structures and mutual interactions with PC lipids. 
\todo{This discussion should be updated.}

>>>>

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPS.eps}
  \caption{\label{HGorderparametersPCvsPS}
    Changes of PC (left panel) and PS (right panel) headgroup order parameters
    from POPC:POPS mixtures with increasing amount of POPS.
    Experimental results of POPC are taken from Ref. \citenum{scherer87}
    (signs are determined as discussed in \cite{botan15,ollila16}).
    Experimental values for POPS in pure bilayer and in mixture are  
    measured in this work and in Ref. \citenum{roux90} at 298K, respectively.
    Since the experimental data of POPS in pure and diluted mixture come from
    different experimental sets (13C NMR in this work and 2H NMR from Ref. \citenum{roux90}),
    the experimental change of the order parameter is less accurate than in typical measurements
    where same technique is used in all conditions, see discussion about qualitative and quantitative 
    accuracy in Ref. \citenum{ollila16}.
    For POPC (left panel) the zero point of y-axis is set to the value of pure bilayer.
    For $\beta$-carbon of POPS (right panel, top) the zero point of y-axis is
    set to the value from POPC:POPS (5:1) mixture.
    For $\alpha$-carbon of POPS (right panel, bottom) the y-axis is tranferred
    with the same value for both order parameters such that the lower order
    parameter value from POPC:POPS (5:1) mixture is at zero
    to correctly illustrate the significant forking.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.} \\
  \todo{Simulation of pure POPC at 298K with Lipid14 would be useful for this plot (only at 303 K is available from NMRlipids I)} \\
  \todo{MacRog simulations of pure POPS with potassium counterions only would be useful for this and other plots.}
\end{figure*}

Oppositely to experiments, the headgroup order parameter of POPS shift away
from zero in CHARM36 simulations when bilayer is diluted with POPC (Fig. \ref{HGorderparametersPCvsPS}).
In lipid14/17 simulations, the POPS order parameter shift closer to zero when
bilayer is diluted with POPC, but the numerical values of order parameters
are too far from experiments to enable interpretation of the experimental data.
Therefore, we conlcude that the force field development is necessary before
MD simulations can be used to interpret the interactions between PC and PS headgroups.
%The $\beta$-carbon order parameter
%increase with increasing amount of PS in CHARMM36 and MacRog simulations in contrast
%to the experimental data.
%The smaller $\alpha$-carbon order parameter increase in both
%simulation models with increasing amount of PS, while it is almost unchanged in experiments.
%The larger $\alpha$-carbon order parameter increase in MacRog and decrease in CHARMM36
%with increasing amount of PS, both model exhibiting a poor agreement with experiments.
